Equalizer 2.0\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Network transport results
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@49 db7c848b-7755-4df2-b763-8796797afa52

\author{Stefan~Eilemann, David~Steiner and Renato Pajarola% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the Visualization and MultiMedia Lab, Department of Informatics, University of Z\"urich.
\IEEEcompsocthanksitem S. Eilemann is also with the Blue Brain Project, Ecole
Polytechnique Federale de Lausanne.}}

@@

%------------------------------------------------------------------------------
\subsection{Infiniband RDMA}

needed? reverse-engr impl

@@

Our RSP implementation uses a separate protocol thread for each RSP connection,
which handles all reads and writes on the multicast group. This thread
implements the protocol handling and communicates with the application threads
through thread-safe queues, as shown in \fig{fRSPProto}. These queues contain
datagrams, which are prefixed by a type-specific header. Each connection has a
configurable number of buffers, (1024 by default) of a configurable MTU (1470
bytes default) which are either free for use or in-use for reading or writing.

%%     1,     // RSP_ERROR_DOWNSCALE
%%     5,     // RSP_ERROR_UPSCALE
%%     20,    // RSP_ERROR_MAXSCALE
%%     3,     // RSP_MIN_SENDRATE_SHIFT

%%     17,    // RSP_ACK_FREQUENCY

@@

Handling a smooth packet flow is critical for performance. RSP uses an activ
flow control to advance the byte stream buffered by the implementation. Each
incoming connection actively acknowledges every $n$ (17 by default) packets
fully received. The incoming connections shift this acknowledgment by their
connection identifier to avoid bursts of acks. Any missed datagram is actively
nack'ed as soon as it is detected. Write connections continuously retransmit
packets from nack datagrams, and advance their \textsf{\_writeBuffers} upon
reception of acks from all readers. Furthermore, the writer will explicitly
request an ack or nack when it runs out of empty buffers or finishes its write
queue.

@@

@@

The spike frequency volumes aggregate the number of spikes which happened within
a given voxel over a given time. The absolute spike count is renormalized to an
unsigned byte value (0..255) during creation. Higher densities in the volume
represent higher spiking activity in the voxel.

@@

The David statue at 2mm resolution is organized in a kD-tree for rendering. Each
node is a separate distributable object, which has two child node objects. A
total of $1023$ objects is distributed and synchronized.

The volume data sets are each distributed in a single object, serializing the
raw volume buffer. The spike volume data set has a significant compression
ratio, which reflects in the results in \fig{fSpikesDist}. Compression for this
data is benefitial for transmitting data over a 10~Gb/s link even for a single
slave process only (970 MB/s vs 570 MB/s raw data transfer). Buffering has
little benefit since the serialization of volume data is trivial. Buffered
compression makes a huge difference, since the compression cost can be amortized
over $n$ nodes, reaching raw data transmission rates of 3.7 GB/s with the
default Snappy compressor and at best 4.4 GB/s with ZStandard at the lowest
compression ratio.

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left) and data synchronization
    times (middle, detail view right) for the spike volume data set in
    \fig{fVolumes}, left}
\end{figure*}

The distribution of the beechnut data set also behaves as expected
(\fig{fBeechnutDist}). Due to the larger object size, uncompressed transmission
is slightly faster compared to the spike data set at 700 MB/s, and compressed
transmission does not improve the mapping performance, likely due to increased
memory pressure caused by the large data size. The comparison of the various
compression engines is consisten with the benchmarks in
\fig{fCompressorDetail}; RLE, Snappy and the LZ variants are very close to each
other, and ZSTD1 can provide better performance after four nodes due to the
better compression ratio.

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/beechnutdist}
  \caption{\label{fBeechnutDist}Object mapping (left) and data synchronization
    times (middle, detail view right) for the beechnut volume data set in
    \fig{fVolumes}, right}
\end{figure*}

\FIXME{Missing: david on 10GE, spike comparisoon 10GE-IB-RSP}

@@

>>>>

\author{Stefan~Eilemann, David~Steiner and Renato Pajarola%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the
Visualization and MultiMedia Lab, Department of Informatics, University of
Z\"urich.
\IEEEcompsocthanksitem S. Eilemann is also with the Blue Brain Project, Ecole
Polytechnique Federale de Lausanne.}}

@@

@@

Our RSP implementation uses a separate protocol thread for each RSP group,
which handles all reads and writes on the multicast group. This thread
implements the protocol handling and communicates with the application threads
through thread-safe queues, as shown in \fig{fRSPProto}. These queues contain
datagrams, which are prefixed by a type-specific header. Each connection has a
configurable number of buffers, (1024 by default) of a configurable MTU (1470
bytes default) which are either free for use or in use for reading or writing.

@@

Handling a smooth packet flow is critical for performance. RSP uses active flow
control to advance the byte stream buffered by the implementation. Each incoming
connection actively acknowledges every $n$ (17 by default) packets fully
received. The incoming connections temporarily offset this acknowledgment by
their connection identifier to avoid bursts of acks. Any missed datagram is
actively nack'ed as soon as it is detected. Write connections continuously
retransmit packets from nack datagrams, and advance their
\textsf{\_writeBuffers} upon reception of acks from all readers. Furthermore,
the writer will explicitly request an ack or nack when it runs out of empty
buffers or finishes its write queue. Nack datagrams may contain multiple ranges
of missed datagrams, as UDP implementations often drop multiple packets at
once.

@@

Bino is a stereoscopic 3D video player capable of running on very large display
systems. Originally written for the immersive semi-cylindrical projection
system at the University of Siegen, it has been used in many installations
thanks to its flexibility of configuration. Bino decodes the video on each
Equalizer rendering process, and only synchronizes the time step globally,
therefore providing a scalable solution to video playback.

@@

The spike frequency volumes aggregate the number of spikes which happened within
a given voxel over a given time. The absolute spike count is renormalized to an
unsigned byte value during creation. Higher densities in the volume represent
higher spiking activity in the voxel.

@@

The David statue at 2mm resolution is organized in a kD-tree for rendering. Each
node is a separate distributed object, which has two child node objects. A
total of $1023$ objects is distributed and synchronized. \fig{fDavidDist} shows
the distribution times for this data set. Due to limited compressibility of the
data, the results are relative similar. Compressing the data repeatedly for
each client leads to decreased performance, since the compression overhead
cannot be amortized by the decreased transmission time. Buffering data slightly
improves performance by reducing the CPU and copy overhead. Combining
compression and buffering leads to the best performance, although only by about
10\%.

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/daviddist}
  \caption{\label{fDavidDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the David 2mm data set in
    \fig{fDavid2mm}}
\end{figure*}

During synchronization, data is pushed from the master process to all mapped
slaves using a unicast connection to each slave. While the results in the
\fig{fDavidDist} (middle) are relatively close to each other, we can still
observe how the tradeoff between compression ratio and speed influences overall
performance. Better, slower compression algorithms lead to better overall better
performance when amortized over many send operations.

The volume data sets are distributed in a single object, serializing the
raw volume buffer. The spike volume data set has a significant compression
ratio, which reflects in the results in \fig{fSpikesDist}. Compression for this
data is benefitial for transmitting data over a 10~Gb/s link even for a single
slave process only (970 MB/s vs 570 MB/s raw data transfer). Buffering has
little benefit since the serialization of volume data is trivial. Buffered
compression makes a huge difference, since the compression cost can be amortized
over $n$ nodes, reaching raw data transmission rates of 3.7 GB/s with the
default Snappy compressor and at best 4.4 GB/s with ZStandard at level 1.

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the spike data set in
    \fig{fVolumes}, left}
\end{figure*}

The distribution of the beechnut data set also behaves as expected
(\fig{fBeechnutDist}). Due to the larger object size, uncompressed transmission
is slightly faster compared to the spike data set at 700 MB/s, and compressed
transmission does not improve the mapping performance, likely due to increased
memory pressure caused by the large data size. The comparison of the various
compression engines is consistent with the benchmarks in
\fig{fCompressorDetail}; RLE, Snappy and the LZ variants are very close to each
other, and ZSTD1 can provide better performance after four nodes due to the
better compression ratio.

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/beechnutdist}
  \caption{\label{fBeechnutDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the beechnut data set in
    \fig{fVolumes}, right}
\end{figure*}

Finally, we compare data distribution speed using different protocols. In this
benchmark, data synchronization time of the spike volume data set is measured,
as in \fig{fSpikesDist} (middle). Buffering is enabled, and compression is
disabled to focus on the raw network performance. \fig{fIFDist} shows the
performance using various protocols.

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/ifdist}
  \caption{\label{fIFDist}Object synchronization using different network
    transports}
\end{figure*}

Unsurprisingly, TCP over the faster InfiniBand link outperforms the cheaper ten
gigabit ethernet link by more than a factor of two. Unexpectedly, the native
RDMA connection performs worse, even though it outperforms IPoIB in a simple
peer-to-peer connection benchmark. This needs further investigations, but we
suspect the abstraction of a byte stream connection chosen by \textsf{Collage}
for historical reasons is not will suited for remote DMA semantics, that is,
one needs to design the network API around zero-copy semantics with managed
memory for modern high-speed transports. Both Infiniband connections show
significant measurement jitter.

Our RSP multicast implementation performs as expected. \textsf{Collage} starts
using multicast to commit new object versions when two or more clients are
mapped, since the transmission to a single client is better done using unicast.
RSP consistently outperforms unicast on the same physical interface and shows
good scaling behavior (2.5 times slower on 16 vs 2 clients on ethernet, 1.8
times slower on InfiniBand). The scaling is significantly better when only one
process per node is used (\fig{fIFDist}, middle: 30\% slower on ethernet, flat
on InfiniBand). The increased transmission time with multiple clients is caused
by a higher probability of packet loss, which increases significantly when using
more than one process per node, as shown in \fig{fIFDist}, right. Infiniband
outperforms ethernet slightly, but is largely limited by the RSP implementation
throughput of preparing and queueing the datagrams to and from the protocol
thread, which we observed in a profiling analysis.

@@

RDMA needs further investigation.
